---
title: "TSA Competition"
author: "Chenjia Liu, Xiyue Zhang, Shubhangi Gupta"
date: "2024-04-22"
subtitle: "https://github.com/vivianzzzzz/ZhangLiuGupta_ENV797_TSA_Competition_S2024.git"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(lubridate)
library(dplyr)
library(forecast) 
library(smooth)
```
#Data Wrangling
We wrangle the load, temperature and humidity data by (1) importing it, (2) converting the date column into a date object using lubridate, (3) calculating the mean load for each day using rowMeans(). For temperature and humidity, we first calculate the hourly values as the average of the value across all workstations, and then use rowmeans() to average the hourly values to daily values. Thus, temperature and humidity daily values are averages of the hourly values across all workstations. (4) checking for NAs, and (5) subsetting the data to only have the columns on the meter_id, date and daily mean. The data for all three variables extends from 1st January 2005 to 30th June 2011.
```{r}
#Load Data
load_raw <- read_excel("./data/load.xlsx")
load <- load_raw %>%
  mutate(date = ymd(date)) %>%
  mutate(daily_mean = rowMeans(select(., 3:26), na.rm = TRUE)) %>%
  filter(!is.na(daily_mean)) %>%
  select(meter_id,date,daily_mean)

#Humidity Data
humidity_raw <- read_excel("./data/relative_humidity.xlsx")

humidity <- humidity_raw %>%
  group_by(date) %>%
  summarise(across(starts_with('rh_ws'), mean))%>%
  mutate(daily_mean = rowMeans(select(., 2:29), na.rm = TRUE)) %>%
  filter(!is.na(daily_mean)) %>%
  select(date,daily_mean)

#Temperature Data
temperature_raw <- read_excel("./data/temperature.xlsx")

temperature <- temperature_raw %>%
  group_by(date) %>%
  summarise(across(starts_with('t_ws'), mean))%>%
  mutate(daily_mean = rowMeans(select(., 2:29), na.rm = TRUE)) %>%
  filter(!is.na(daily_mean)) %>%
  select(date,daily_mean)

temperature <- temperature[-2373,]
```

#Creating a time series, training and testing data 
We convert the load, temperature and humidity daily mean datasets into time series objects using hte msts() function. Seasonal periods are taken to be 365.25. The training data extends from 1st January 2005 to 31st May 2011 and the testing data extends from 1st-30th June 2011.
```{r}
#Load Data
ts_load <- msts(load$daily_mean, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

ts_load_training <- subset(ts_load,end = length(ts_load)-30)
ts_load_testing <- subset(ts_load,start = length(ts_load)-30)

ts_load_training %>% mstl() %>%
  autoplot()

#Temperature Data
ts_temperature <- msts(temperature$daily_mean, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

ts_temperature_training <- subset(ts_load,end = length(ts_load)-30)
ts_temperature_testing <- subset(ts_load,start = length(ts_load)-30)

#Humidity Data
ts_humidity <- msts(humidity$daily_mean, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

ts_humidity_training <- subset(ts_load,end = length(ts_load)-30)
ts_humidity_testing <- subset(ts_load,start = length(ts_load)-30)

```
#Fitting models to the training data and forecasting on testing data
In this section, we fit the following models to the load training data and forecast them for the next 30 days that make up the testing data: (1) STL + ETS (2) TBATS (3) 7 versions of Neural Networks - 4 using different combinations of p and P, 1 with temperature, 1 with humidity, and 1 with temperature+humidity (4) 2 versions of ARIMA - one with temperature, and one with humidity, and (5) SS Exponential smoothing
The results of this data have been printed in a table at the end of this section.

##Fitting models only on load data
```{r}
#LOAD ONLY
#(1) STL + ETS model
ETS_fit <-  stlf(ts_load_training,h=30)
autoplot(ETS_fit)
ETS_scores <- accuracy(ETS_fit$mean,ts_load_testing)
print(ETS_scores)
```


```{r}
# (2) TBATS model
TBATS_fit <- tbats(ts_load_training)

TBATS_forcast <- forecast(TBATS_fit, h=30)

autoplot(ts_load_testing) +
  autolayer(TBATS_forcast, series="TBATS",PI=FALSE)

TBATS_scores <- accuracy(TBATS_forcast$mean,ts_load_testing)

print(TBATS_scores)
```
```{r}
# (3.1) Neural Network (p=1, P=1)
NN_fit1 <- nnetar(ts_load_training,p=1,P=1,xreg=fourier(ts_load_training, K=c(2,12)))
NN_for1 <- forecast(NN_fit1,h=30, xreg=fourier(ts_load_training, K=c(2,12),h=30))

autoplot(ts_load_testing) +
  autolayer(NN_for1, series="Neural Network",PI=FALSE)

NN_scores1 <- accuracy(NN_for1$mean,ts_load_testing)

print(NN_scores1)
```


```{r}
#(3.2) Neural Network (p=2, P=2)
NN_fit2 <- nnetar(ts_load_training,p=2,P=1,xreg=fourier(ts_load_training, K=c(2,12)))
NN_for2 <- forecast(NN_fit2,h=30, xreg=fourier(ts_load_training, K=c(2,12),h=30))

autoplot(ts_load_testing) +
  autolayer(NN_for2, series="Neural Network",PI=FALSE)

NN_scores2 <- accuracy(NN_for2$mean,ts_load_testing)

print(NN_scores2)
```

```{r}
#(3.3) Neural Network (p=3, P=1)
NN_fit3 <- nnetar(ts_load_training,p=3,P=1,xreg=fourier(ts_load_training, K=c(2,12)))
NN_for3 <- forecast(NN_fit3,h=30, xreg=fourier(ts_load_training, K=c(2,12),h=30))

autoplot(ts_load_testing) +
  autolayer(NN_for3, series="Neural Network",PI=FALSE)

NN_scores3 <- accuracy(NN_for3$mean,ts_load_testing)

print(NN_scores3)
```

```{r}
#(3.4) Neural Network (p=1, P=0)
NN_fit4 <- nnetar(ts_load_training,p=1,P=0,xreg=fourier(ts_load_training, K=c(2,12)))
NN_for4 <- forecast(NN_fit4,h=30, xreg=fourier(ts_load_training, K=c(2,12),h=30))

autoplot(ts_load_testing) +
  autolayer(NN_for4, series="Neural Network",PI=FALSE)

NN_scores4 <- accuracy(NN_for4$mean,ts_load_testing)

print(NN_scores4)
```
##Forecasting with temperature and humidity
```{r}
#With temperature data
temp_regressors<- as.matrix(data.frame(fourier(ts_load_training, K=c(2,12)), "temp"= ts_temperature_training))

temp_for<-forecast(ts_temperature_training,h=30)

temp_regressors_for<-as.matrix(data.frame(fourier(ts_load_training, K=c(2,12),h=30), "temp"= temp_for$mean))

#with humidity data
hum_regressors<- as.matrix(data.frame(fourier(ts_load_training, K=c(2,12)), "hum"= ts_humidity_training))

hum_for<-forecast(ts_humidity_training,h=30)

hum_regressors_for<-as.matrix(data.frame(fourier(ts_load_training, K=c(2,12),h=30), "hum"= hum_for$mean))

#With both temperature and humidity
temp_hum_regressors<- as.matrix(data.frame(fourier(ts_load_training, K=c(2,12)), "temp"= ts_temperature_training, "hum"= ts_humidity_training))

temp_hum_regressors_for<-as.matrix(data.frame(fourier(ts_load_training, K=c(2,12),h=30), "temp"= temp_for$mean,"hum"= hum_for$mean))
```


```{r}
# (3.5) Neural Network + Temperature
NN_fit5 <- nnetar(ts_load_training,p=1,P=0,xreg=temp_regressors)
NN_for5 <- forecast(NN_fit5,h=30, xreg=temp_regressors_for)

autoplot(ts_load_testing) +
  autolayer(NN_for5, series="Neural Network",PI=FALSE)

NN_scores5 <- accuracy(NN_for5$mean,ts_load_testing)

print(NN_scores5)

```

```{r}
# (3.6) Neural Network + Humidity
NN_fit6 <- nnetar(ts_load_training,p=1,P=0,xreg=hum_regressors)
NN_for6 <- forecast(NN_fit6,h=30, xreg=hum_regressors_for)

autoplot(ts_load_testing) +
  autolayer(NN_for6, series="Neural Network",PI=FALSE)

NN_scores6 <- accuracy(NN_for6$mean,ts_load_testing)

print(NN_scores6)
```

```{r}
# (3.7) Neural Network + Temperature + Humidity
NN_fit7 <- nnetar(ts_load_training,p=1,P=0,xreg=temp_hum_regressors)
NN_for7 <- forecast(NN_fit7,h=30, xreg=temp_hum_regressors_for)

autoplot(ts_load_testing) +
  autolayer(NN_for7, series="Neural Network",PI=FALSE)

NN_scores7 <- accuracy(NN_for7$mean,ts_load_testing)

print(NN_scores7)
```
```{r}
# (4.1) Arima + Temperature
ARIMA_fit1<-auto.arima(ts_load_training,seasonal= FALSE, lambda=0,xreg=temp_regressors)
ARIMA_for1<-forecast(ARIMA_fit1,xreg=temp_regressors_for,h=30)

autoplot(ts_load_testing) +
  autolayer(ARIMA_for1, series="ARIMA",PI=FALSE)

ARIMA_scores1 <- accuracy(ARIMA_for1$mean,ts_load_testing)

print(ARIMA_scores1)
```

```{r}
# (4.2) Arima + humidity
ARIMA_fit2<-auto.arima(ts_load_training,seasonal= FALSE, lambda=0,xreg=hum_regressors)
ARIMA_for2<-forecast(ARIMA_fit2,xreg=hum_regressors_for,h=30)

autoplot(ts_load_testing) +
  autolayer(ARIMA_for2, series="ARIMA",PI=FALSE)

ARIMA_scores2 <- accuracy(ARIMA_for2$mean,ts_load_testing)

print(ARIMA_scores2)
```

```{r}
#(5) SS Exponential smoothing

SSES_fit1 <- es(ts_load_training,model="ZZZ",h=30,holdout=FALSE)
plot(SSES_fit1)
SSES_scores1 <- accuracy(SSES_fit1$forecast,ts_load_testing)
print(SSES_scores1)
```

##Printing the scores of all the above fitted models in one table
```{r}
scores <- as.data.frame(
  rbind(TBATS_scores, NN_scores1,NN_scores2, NN_scores3,NN_scores4, NN_scores5, NN_scores6, ARIMA_scores1)
  )
row.names(scores) <- c("TBATS_scores", "NN_scores1","NN_scores2","NN_scores3","NN_scores4","NN_scores5","NN_scores6", "ARIMA_scores1")
scores
```

#Forecasting daily demand for July 2011
In this section, we use the fitted models to forecast daily demand for July 2011. These results have been uploaded on Kaggle. We forecast the following models: (1) SS Exponential Smoothing (2) TBATS (3) 5 Neural Network Models - 2 with different combinations of p and P, one with temperature, one with humidity, and one with temperature and humidity (4) two ARIMA models - one with temperature, and one with humidity

##Forecasting only load data

```{r}
# (1) SS Exponential smoothing
SSES_fit_load <- es(ts_load,model="ZZZ",h=31,holdout=FALSE)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-SSES_fit_load$forecast
July_SSES <-data.frame(date=date, load=load)
July_SSES
write.csv(July_SSES, file = "July_SSES.csv", row.names = FALSE)
```
```{r}
# (2) TBATS
TBATS_fit_load <- tbats(ts_load)
TBATS_forcast_load <- forecast(TBATS_fit_load, h=31)

autoplot(ts_load) +
  autolayer(TBATS_forcast_load, series="TBATS",PI=FALSE)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-TBATS_forcast_load$mean
July_TBATS<-data.frame(date=date, load=load)
July_TBATS
write.csv(July_TBATS, file = "July_TBATS.csv", row.names = FALSE)
```

```{r}
# (3.1) Neural Network (p=1, P=1)
NN_fit1_load <- nnetar(ts_load,p=1,P=1,xreg=fourier(ts_load, K=c(2,12)))
NN_for1_load <- forecast(NN_fit1_load,h=31, xreg=fourier(ts_load, K=c(2,12),h=31))

autoplot(ts_load) +
  autolayer(NN_for1_load, series="Neural Network",PI=FALSE)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-NN_for1_load$mean
July_NN1<-data.frame(date=date, load=load)
July_NN1
write.csv(July_NN1, file = "July_NN1.csv", row.names = FALSE)
```

```{r}
# (3.2) Neural Network (p=1, P=0)
NN_fit4_load <- nnetar(ts_load,p=1,P=0,xreg=fourier(ts_load, K=c(2,12)))
NN_for4_load <- forecast(NN_fit4_load,h=31, xreg=fourier(ts_load, K=c(2,12),h=31))

autoplot(ts_load) +
  autolayer(NN_for4_load, series="Neural Network",PI=FALSE)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-NN_for4_load$mean
July_NN4<-data.frame(date=date, load=load)
July_NN4
write.csv(July_NN4, file = "July_NN4.csv", row.names = FALSE)
```

##Forecasting load data with regressors for temperature and humidity
```{r}
#Temperature
temp_regressors_load<- as.matrix(data.frame(fourier(ts_load, K=c(2,12)), "temp"= ts_temperature))

temp_for_load<-forecast(ts_temperature,h=31)
temp_regressors_for_load<-as.matrix(data.frame(fourier(ts_load, K=c(2,12),h=31), "temp"= temp_for_load$mean))

#Humidity

hum_regressors_load<- as.matrix(data.frame(fourier(ts_load, K=c(2,12)), "hum"= ts_humidity))

hum_for_load<-forecast(ts_humidity,h=31)
hum_regressors_for_load<-as.matrix(data.frame(fourier(ts_load, K=c(2,12),h=31), "hum"= hum_for_load$mean))

#Temperature & Humidity
temp_hum_regressors_load<- as.matrix(data.frame(fourier(ts_load, K=c(2,12)), "temp"= ts_temperature, "hum"= ts_humidity))

temp_hum_regressors_for_load<-as.matrix(data.frame(fourier(ts_load, K=c(2,12),h=31), "temp"= temp_for_load$mean,"hum"= hum_for_load$mean))
```

```{r}
# (3.3) Neural Network + Temperature
NN_fit5_load <- nnetar(ts_load,p=1,P=0,xreg=temp_regressors_load)
NN_for5_load <- forecast(NN_fit5_load,h=31, xreg=temp_regressors_for_load)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<- NN_for5_load$mean
July_NN5 <-data.frame(date=date, load=load)
write.csv(July_NN5, file = "July_NN5.csv", row.names = FALSE)
```

```{r}
# (3.4) Neural Network + Humidity
NN_fit6_load <- nnetar(ts_load,p=1,P=0,xreg=hum_regressors_load)
NN_for6_load <- forecast(NN_fit6_load,h=31, xreg=hum_regressors_for_load)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-NN_for6_load$mean
July_NN6 <-data.frame(date=date, load=load)
write.csv(July_NN6, file = "July_NN6.csv", row.names = FALSE)
```

```{r}
# (3.5) Neural Network + Temperature + Humidity
NN_fit7_load <- nnetar(ts_load,p=1,P=0,xreg=temp_hum_regressors_load)
NN_for7_load <- forecast(NN_fit7_load,h=31, xreg=temp_hum_regressors_for_load)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-NN_for7_load$mean
July_NN7 <-data.frame(date=date, load=load)
write.csv(July_NN7, file = "July_NN7.csv", row.names = FALSE)
```

```{r}
# (4.1) Arima+Temperature
ARIMA_fit1_load<-auto.arima(ts_load,seasonal= FALSE, lambda=0,xreg=temp_regressors_load)
ARIMA_for1_load<-forecast(ARIMA_fit1_load,xreg=temp_regressors_for_load,h=31)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-ARIMA_for1_load$mean
ARIMA_for1_load <-data.frame(date=date, load=load)
write.csv(ARIMA_for1_load, file = "July_ARIMA1.csv", row.names = FALSE)
```

```{r}
# (4.2) Arima+humidity
ARIMA_fit2_load<-auto.arima(ts_load,seasonal= FALSE, lambda=0,xreg=hum_regressors_load)
ARIMA_for2_load<-forecast(ARIMA_fit2_load,xreg=hum_regressors_for_load,h=31)

#Exporting into a CSV
date <- seq(ymd("2011-07-01"), ymd("2011-07-31"), by = "days")
load<-ARIMA_for2_load$mean
ARIMA_fit2_load <-data.frame(date=date, load=load)
write.csv(ARIMA_fit2_load, file = "July_ARIMA2.csv", row.names = FALSE)

```

